---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
- Load data
````{r echo=TRUE}
unzip("activity.zip")
#rawData <- read.csv("activity.csv", header = TRUE, sep = ",")
rawData <- read.csv("activity.csv")
````
- Process/transform the data (if necessary) into a format suitable for your analysis
```{r echo=TRUE}
sumSteps <- aggregate(steps ~ date, data = rawData, sum, na.rm = TRUE)
````


## What is mean total number of steps taken per day?
- Make a histogram of the total number of steps taken each day
```{r echo=TRUE}
hist(sumSteps$steps,col="green",xlab="Number of steps per day",main="Histogram of number of steps per day")
```

- Calculate and report the mean and median of the total number of steps taken per day

mean of total number of steps per day
```{r echo=TRUE}
# mean
ms <- mean(sumSteps$steps)
ms
```

median of total number of steps per day
```{r echo=TRUE}
# median
median(sumSteps$steps)
```

## What is the average daily activity pattern?
- Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r echo=TRUE}
stepInterv <- aggregate(steps ~ interval,data=rawData,mean,na.rm=TRUE)
plot(steps ~ interval,data = stepInterv,type="l")
```

- Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r echo=TRUE}
maxInterval <- stepInterv[stepInterv$steps==max(stepInterv$steps),]
maxInterval$interval
```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
